An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures

### 2021-05-27 CMPSpaper_writing/CMPSpaper_writing.Rmd

---
title: An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures
header-includes:
  - \usepackage{xcolor}
  - \usepackage{longtable}
  - \usepackage{adjustbox}            
  - \usepackage[font=small,skip=5pt]{caption}
  - \usepackage{subcaption}
  - \usepackage{afterpage}
author:
  # see ?rjournal_article for more information
  - name: Will W. Ju
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
    url: https://github.com/willju-wangqian
    orcid: 0000-0002-9079-593X
    email:  wju@iastate.edu
  - name: Heike Hofmann
    url: https://github.com/heike
    email: hofmann@iastate.edu
    orcid: 0000-0002-9079-593X
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble

@@

1. **cut the reference signature into consecutive, non-overlapping and equal-length basis segments:**
The command `get_segs(x, len=50)` implements this step:  it takes  bullet signature `x` in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`, which are referred as "basis segments".

@@

2. **identify candidate positions:** For each basis segment a set of candidate \wj{registration XXX} positions on the comparison signature is identified based on the segment's similarity to the comparison. In a first step, the cross-correlation function of the segment to the comparison is calculated, then a number of positions with high correlation values are identified as candidate positions. In case multiple segment lengths are considered, the length of each basis segment is doubled and these two steps are repeated. Only when candidate positions coincide (or are similar enough), they are considered further. \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.

@@

    +  **Multiple-length segments:** If `npeaks.set` is an integer vector of length more than 1, for example `npeaks.set = c(5, 3, 1)`, a multi segment lengths strategy will be used: finding top five peaks in the original scale, double the segment length and recalculate the cross-correlation function, then identify 3 peaks in the second scale and repeat the process for a basis segment that is doubled again to identify 1 peak in the third scale. \autoref{fig:seg_all} shows an example of 3 scales of basis segment 6 and their corresponding cross-correlation curves and identified peaks. If a single position (with a tolerance zone determined by `Tx`) results in a top peak in the cross-correlation curve in all three scales, it is called a consistent correlation peak and identified as a candidate for registration. Note that in @cmps the last element of `npeaks.set` in the multi segment lengths strategy is always set to be 1. Additionally, this strategy is also referred as "multi-peak inspection at different segment lengths" in @cmps.

3. **determine the congruent registration position:** Each candidate position corresponding to each of the basis segments gets one `vote`. Votes for candidate positions within a tolerance of `Tx` are tallied and the position with the highest number of votes gets chosen as the *congruent registration position*. Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score of the comparison.

@@

* `get_ccp(ccr.list, Tx = 25)` tries to identify the "consistent correlation peak" (defined in algorithm step 2) under the framework of multi segment lengths strategy. `ccr.list` is described in the previous bullet point, and `Tx` determines the size of a tolerance zone used in identifying the consistent correlation peak.

* `get_CMPS(input.ccp, Tx = 25)` tallies votes from all basis segments and determines the congruent registration position. A position "receives" a vote from a basis segment if this basis segment identifies this position or a closed-enough position (within a tolerance zone defined by `Tx`) as a candidate registration position in algorithm step 2. The goal is to find a shift in registration position that is agreed by the most number of basis segments, so the position receiving the most number of votes is the congruent registration position. In case of ties the middle position is taken as the congruent registration position. Basis segments with the congruent registration position identified are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score.

>>>>

---
title: An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures
header-includes:
  - \usepackage{xcolor}
  - \usepackage{longtable}
  - \usepackage{adjustbox}            
  - \usepackage[font=small,skip=5pt]{caption}
  - \usepackage{subcaption}
  - \usepackage{afterpage}
author:
  # see ?rjournal_article for more information
  - name: Will W. Ju
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
    url: https://github.com/willju-wangqian
    email:  wju@iastate.edu
  - name: Heike Hofmann
    url: https://github.com/heike
    email: hofmann@iastate.edu
    orcid: 0000-0002-9079-593X
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
abstract: >
  An abstract of less than 150 words.
preamble: |
  % Any extra LaTeX you need in the preamble

@@

1. **cut the reference signature into consecutive, non-overlapping and equal-length basis segments:**
The command `get_segs(x, len=50)` implements this step:  it takes  bullet signature `x` in the format of a numeric vector and cuts it into consecutive, non-overlapping and equal-length segments of length `len`, which are referred to as "basis segments".

@@

2. **identify candidate positions:** For each basis segment a set of candidate registration  positions on the comparison signature is identified based on the segment's similarity to the comparison. In a first step, the cross-correlation function of the segment to the comparison is calculated, then a number of positions with high correlation values are identified as candidate positions. In case multiple segment lengths are considered, the length of each basis segment is doubled and these two steps are repeated. Only when candidate positions coincide (or are similar enough), they are considered further. \autoref{fig:segplots} and \autoref{fig:seg_all} illustrate these ideas.

@@

    +  **Multiple-length segments:** If `npeaks.set` is an integer vector of length more than 1, for example `npeaks.set = c(5, 3, 1)`, a multi segment lengths strategy will be used: finding top five peaks in the original scale, double the segment length and recalculate the cross-correlation function, then identify 3 peaks in the second scale and repeat the process for a basis segment that is doubled again to identify 1 peak in the third scale. \autoref{fig:seg_all} shows an example of 3 scales of basis segment 6 and their corresponding cross-correlation curves and identified peaks. If a single position (with a tolerance zone determined by `Tx`) results in a top peak in the cross-correlation curve in all three scales, it is called a consistent correlation peak and identified as a candidate position for registration. Note that in @cmps the last element of `npeaks.set` in the multi segment lengths strategy is always set to be 1. Additionally, this strategy is also referred as "multi-peak inspection at different segment lengths" in @cmps.

3. **determine the congruent registration position:** Each candidate position corresponding to each of the basis segments gets one `vote`. Votes for candidate positions within a tolerance of `Tx` are tallied and the position with the highest number of votes gets chosen as the *congruent registration position*. Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score of the comparison. \hh{XXX what happens when there are ties after the vote?}

@@

* `get_ccp(ccr.list, Tx = 25)` tries to identify the "consistent correlation peak" (defined in algorithm step 2) under the framework of multi segment lengths strategy. `ccr.list` is described in the previous bullet point, and `Tx` determines the size of a tolerance zone used in identifying the consistent correlation peak. \hh{XXX what happens when there is no consistent correlation peak?}

* `get_CMPS(input.ccp, Tx = 25)` \hh{XXX how is `get\_CMPS` different from the extract function?} tallies votes from all basis segments and determines the congruent registration position. A position "receives" a vote from a basis segment if this basis segment identifies this position or a closed-enough position (within a tolerance zone defined by `Tx`) as a candidate registration position in algorithm step 2. The goal is to find a shift in registration position that is agreed by the most number of basis segments, so the position receiving the most number of votes is the congruent registration position. In case of ties the middle position is taken as the congruent registration position. Basis segments with the congruent registration position identified are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score.
